---
title: "Pre-analysis: Psychological functioning across the pandemic: A comparison of Canadian young adults with and without chronic illness"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load the libraries
library(tidyverse)
library(readr)
library(here)
library(psych)
library(english)
library(stringr)
library(tableone)
library(kableExtra)
library(scales)
library(ggridges)
library(plotly)
library(patchwork)
library(haven)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(table1)
library(rstatix)
library(car)
library(multcomp)
#library(compute.se)
library(effects)
library(pastecs)
#library(WRS)
```

```{r data, echo=FALSE, message=FALSE, warning=FALSE}
# Read in the clean csv file
leger <- read_csv("data/cmaj_analyses_jan2022.csv")

# subset for chronic people only
leger_chronic <- leger %>%
  filter(chronic_dummy == 1)

# Data for the plot
marginal_means <- read_csv("data/marginal_means.csv")

marginal_means_2 <- read_csv("data/marginal_means2.csv")

covid_cases <- read_csv("data/covid19-download.csv")

# set seed for all analyses and wrangling
set.seed(1)
```

#### Aims:

This document contains some of the exploratory code that led to the final analyses. The exact steps are outlined in the "project log".

The present study is part of a larger International COVID‐19 Awareness and Responses Evaluation (iCARE) study, which is a global multi‐phase research survey conducted through the Montreal Behavioural Medicine Centre. Find more about the study and materials at [www.iCAREstudy.com](www.iCAREstudy.com). 

### COVID-19 related psychological distress over time

```{r echo=FALSE, warning=FALSE, fig.width= 12, fig.height=6}
leger %>%
  dplyr::select(chronic_dummy, wave, distress_sum, sex) %>%
  filter(sex %in% c("Female", "Male")) %>%
  mutate(chronic_dummy = factor(chronic_dummy,
                                levels = c(0:1),
                                labels = c("No", "Yes")),
         wave = as.factor(wave)) %>%
  ggplot(aes(x = wave)) +
  geom_boxplot(aes(y = distress_sum, fill=chronic_dummy)) +
  #facet_wrap(~chronic_dummy) +
  labs(title = " ", x = "Survey round", y = "Level of psychological distress", fill = "Chronic condition") +
  #facet_grid(~sex) +
  ylim(0, 13) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_discrete(breaks = c(2, 3, 4, 5, 6, 7),
                     labels = c("Mid June \n2020 \n(2)", "Early Nov. \n2020 \n(3)", "Late Jan. \n2021 \n(4)",
                                "Late March \n2021 \n(5)", "Early June \n2021 \n(6)", "Mid Sept. \n2021 \n(7)"))


GeomSplitViolin <- ggproto("GeomSplitViolin", GeomViolin, 
                           draw_group = function(self, data, ..., draw_quantiles = NULL) {
  data <- transform(data, xminv = x - violinwidth * (x - xmin), xmaxv = x + violinwidth * (xmax - x))
  grp <- data[1, "group"]
  newdata <- plyr::arrange(transform(data, x = if (grp %% 2 == 1) xminv else xmaxv), if (grp %% 2 == 1) y else -y)
  newdata <- rbind(newdata[1, ], newdata, newdata[nrow(newdata), ], newdata[1, ])
  newdata[c(1, nrow(newdata) - 1, nrow(newdata)), "x"] <- round(newdata[1, "x"])

  if (length(draw_quantiles) > 0 & !scales::zero_range(range(data$y))) {
    stopifnot(all(draw_quantiles >= 0), all(draw_quantiles <=
      1))
    quantiles <- ggplot2:::create_quantile_segment_frame(data, draw_quantiles)
    aesthetics <- data[rep(1, nrow(quantiles)), setdiff(names(data), c("x", "y")), drop = FALSE]
    aesthetics$alpha <- rep(1, nrow(quantiles))
    both <- cbind(quantiles, aesthetics)
    quantile_grob <- GeomPath$draw_panel(both, ...)
    ggplot2:::ggname("geom_split_violin", grid::grobTree(GeomPolygon$draw_panel(newdata, ...), quantile_grob))
  }
  else {
    ggplot2:::ggname("geom_split_violin", GeomPolygon$draw_panel(newdata, ...))
  }
})

geom_split_violin <- function(mapping = NULL, data = NULL, stat = "ydensity", position = "identity", ..., 
                              draw_quantiles = NULL, trim = TRUE, scale = "area", na.rm = FALSE, 
                              show.legend = NA, inherit.aes = TRUE) {
  layer(data = data, mapping = mapping, stat = stat, geom = GeomSplitViolin, 
        position = position, show.legend = show.legend, inherit.aes = inherit.aes, 
        params = list(trim = trim, scale = scale, draw_quantiles = draw_quantiles, na.rm = na.rm, ...))
}

set.seed(20160229)

leger %>%
  dplyr::select(chronic_dummy, wave, distress_sum, sex) %>%
  filter(sex %in% c("Female", "Male")) %>%
  mutate(m = factor(chronic_dummy,
                                levels = c(0:1),
                                labels = c("No", "Yes")),
         x = as.factor(wave),
         y = distress_sum) %>%
  ggplot(aes(x, y, fill = m)) +
  geom_split_violin() +
  labs(title = "Split violin (density) plot", x = "Survey round", y = "Level of psychological distress", fill = "Chronic condition") +
  #facet_grid(~sex) +
  ylim(0, 12) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_discrete(breaks = c(2, 3, 4, 5, 6, 7),
                     labels = c("Mid June \n2020 \n(2)", "Early Nov. \n2020 \n(3)", "Late Jan. \n2021 \n(4)",
                                "Late March \n2021 \n(5)", "Early June \n2021 \n(6)", "Mid Sept. \n2021 \n(7)"))

###
leger %>%
  dplyr::select(chronic_dummy, wave, distress_sum, sex) %>%
  filter(sex %in% c("Female", "Male")) %>%
  mutate(z = -log(distress_sum+1)) %>%
  mutate(chronic_dummy = factor(chronic_dummy,
                                levels = c(0:1),
                                labels = c("No", "Yes")),
         wave = as.factor(wave)) %>%
  ggplot(aes(x = z)) +
  geom_histogram(aes(fill = z), bins = 10) +
  facet_grid(wave~chronic_dummy)

  geom_boxplot(aes(y = distress_sum, fill=chronic_dummy)) +
  #facet_wrap(~chronic_dummy) +
  labs(title = " ", x = "Survey round", y = "Level of psychological distress", fill = "Chronic condition") +
  #facet_grid(~sex) +
  ylim(0, 13) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_discrete(breaks = c(2, 3, 4, 5, 6, 7),
                     labels = c("Mid June \n2020 \n(2)", "Early Nov. \n2020 \n(3)", "Late Jan. \n2021 \n(4)",
                                "Late March \n2021 \n(5)", "Early June \n2021 \n(6)", "Mid Sept. \n2021 \n(7)"))


#test <- 
leger %>%
  #dplyr::select(distress_sum)
  #mutate(z = scale(distress_sum)) %>%
  dplyr::select(distress_sum, chronic_dummy, wave) %>%
  group_by(wave, chronic_dummy) %>%
  shapiro_test(chronic_dummy)
```

### Density plots between the health status groups and age bracket

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width= 12, fig.height=6}
leger %>%
  mutate(age_grp = factor(case_when(age_yrs <26 ~ "Under 25 years",
                                    age_yrs >25 ~ "26-39 years"),
                          levels = c("Under 25 years", "26-39 years"),
                          ordered = TRUE),
         chronic_dummy = factor(chronic_dummy,
                                levels = c(0:1),
                                labels = c("Healthy", "Chronic health"))) %>%
ggplot(aes(distress_sum, fill=chronic_dummy)) +
  geom_density(alpha=.3) +
  scale_fill_viridis_d(option = "D") +
  theme_bw(base_size = 15) +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major = element_blank(),
        strip.text = element_text(face = 'bold')) +
  labs(x = "Distress", y = "Density", fill = " ") +
  facet_grid(~age_grp)


leger %>%
  mutate(age_grp = factor(case_when(age_yrs <26 ~ "Under 25 years",
                                    age_yrs >25 ~ "26-39 years"),
                          levels = c("Under 25 years", "26-39 years"),
                          ordered = TRUE),
         chronic_dummy = factor(chronic_dummy,
                                levels = c(0:1),
                                labels = c("Healthy", "Chronic health"))) %>%
  filter(sex %in% c("Female", "Male")) %>%
  group_by(age_grp, chronic_dummy) %>%
  count() %>%
  drop_na() %>%
  group_by(age_grp) %>%
  mutate(sum_grp = sum(n),
         prop = round((n/sum_grp*100),2)) %>%
  dplyr::select(-sum_grp) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

### FEMALE v MALE v CHRONIC v HEALTHY
```{r echo=FALSE, warning=FALSE, message=FALSE}
test1 <- leger %>%
  dplyr::select(age_yrs, sex, chronic_dummy, distress_sum, wave) %>%
  filter(sex %in% c("Female", "Male")) %>%
  mutate(age_grp = factor(case_when(age_yrs <26 ~ "Under 25 years",
                                    age_yrs >25 ~ "26-39 years"),
                          levels = c("Under 25 years", "26-39 years"),
                          ordered = TRUE),
         chronic_dummy = factor(chronic_dummy,
                                levels = c(0:1),
                                labels = c("Healthy", "Chronic health")),
         wave = factor(wave,ordered = TRUE))

test1 %>%
  group_by(age_grp, chronic_dummy, sex) %>%
  summarise(means = mean(distress_sum, na.rm = TRUE)) %>%
  ggplot(aes(x = age_grp, y = means, color = chronic_dummy, group = chronic_dummy)) +
  geom_point(size = 4) +
  geom_line() +
  facet_wrap(~sex) +
  theme_bw(base_size = 15) +
  theme(strip.text = element_text(face = 'bold')) +
  labs(x = "Age group", y = "Distress level", fill = " ") +
  ylim(0, 12)
```

<!-- ### Associations between psych distress and socio-demographic factors -->

<!-- #### Logistic regression output -->

<!-- *NOTE* - Education is treated as continuous variable with the following levels: "never been" ~ 0, "primary" ~ 1,"secondary" ~ 2, "college" ~ 3, "graduate" ~ 4. -->

<!-- Province is broken down into 3 geographical regions:  -->

<!-- * *East* = "New Brunswick", "Nova Scotia", "Prince Edward Island", "Newfoundland" -->
<!-- * *Central* = "Ontario", "Quebec" -->
<!-- * *West* = "British Columbia", "Alberta", "Saskatchewan", "Manitoba" -->

<!-- Distress rev = compound score of COVID-19 related feelings of anxiety, depression, and loneliness. Scale rangess from 4 - 16, with higher scores indicating higher levels of pandemic-related distress. -->

<!-- *** -->

<!-- ## CHRONIC GROUP -->

```{r echo=FALSE, warning=FALSE}
# set.seed(1)
# 
# leger_chronic_m1 <- leger_chronic %>%
#   select(vaccine_dummy, age_yrs, sex, hoinc, edu, province, wave, any_m_health, chronic_multi, distress, concern_sq002) %>%
#   filter(sex %in% c("Male", "Female")) %>%
#   mutate(education = as.numeric(case_when(edu == "Never been" ~ 0, 
#                                           edu == "Primary" ~ 1,
#                                           edu == "Secondary" ~ 2,
#                                           edu == "College" ~ 3,
#                                           edu == "Graduate" ~ 4)))
# 
# m1_chronic <- glm(distress ~ age_yrs + sex + province + education + hoinc + any_m_health + chronic_multi + wave + concern_sq002,
#                   data = leger_chronic_m1)
# 
# tab_model(m1_chronic)

```

<!-- ## HEALTHY GROUP -->
```{r echo=FALSE, warning=FALSE}
# set.seed(1)
# 
# leger_healthy_m1 <- leger %>%
#   filter(chronic_dummy == 0) %>%
#   select(vaccine_dummy, age_yrs, sex, hoinc, edu, province, wave, any_m_health, chronic_multi, distress, concern_sq002) %>%
#   filter(sex %in% c("Male", "Female")) %>%
#   mutate(education = as.numeric(case_when(edu == "Never been" ~ 0, 
#                                           edu == "Primary" ~ 1,
#                                           edu == "Secondary" ~ 2,
#                                           edu == "College" ~ 3,
#                                           edu == "Graduate" ~ 4)))
# 
# m1_healthy <- glm(distress ~ age_yrs + sex + province + education + hoinc + any_m_health + wave + concern_sq002,
#                   data = leger_healthy_m1)
# 
# tab_model(m1_healthy)
```


### Plotting Covid cases in Canada 

The first figure was created in hopes to reproduce a figure that someone else in the team has previously come up with - it uses cases summed up by month.

The original figure can be viewed [here](https://www.medrxiv.org/content/medrxiv/early/2021/11/11/2021.11.10.21266174.full.pdf), at the end of the pdf.

```{r echo = FALSE, warning = FALSE, message = FALSE}
covid_cases_test <- read_csv("data/covid19-download.csv") %>%
  dplyr::select(date, prname, numtoday) %>%
  mutate(month = case_when(date >= "2020-01-01" & date < "2020-02-01" ~ 202001,
                           date >= "2020-02-01" & date < "2020-03-01" ~ 202002,
                           date >= "2020-03-01" & date < "2020-04-01" ~ 202003,
                           date >= "2020-04-01" & date < "2020-05-01" ~ 202004,
                           date >= "2020-05-01" & date < "2020-06-01" ~ 202005,
                           date >= "2020-06-01" & date < "2020-07-01" ~ 202006,
                           date >= "2020-07-01" & date < "2020-08-01" ~ 202007,
                           date >= "2020-08-01" & date < "2020-09-01" ~ 202008,
                           date >= "2020-09-01" & date < "2020-10-01" ~ 202009,
                           date >= "2020-10-01" & date < "2020-11-01" ~ 202010,
                           date >= "2020-11-01" & date < "2020-12-01" ~ 202011,
                           date >= "2020-12-01" & date < "2021-01-01" ~ 202012,
                           date >= "2021-01-01" & date < "2021-02-01" ~ 202101,
                           date >= "2021-02-01" & date < "2021-03-01" ~ 202102,
                           date >= "2021-03-01" & date < "2021-04-01" ~ 202103,
                           date >= "2021-04-01" & date < "2021-05-01" ~ 202104,
                           date >= "2021-05-01" & date < "2021-06-01" ~ 202105,
                           date >= "2021-06-01" & date < "2021-07-01" ~ 202106,
                           date >= "2021-07-01" & date < "2021-08-01" ~ 202107,
                           date >= "2021-08-01" & date < "2021-09-01" ~ 202108,
                           date >= "2021-09-01" & date < "2021-10-01" ~ 202109)) %>%
  dplyr::select(-date)

# covid_cases_test %>%
#   filter(prname == "Canada") %>%
#   filter(month == 202105) %>%
#   mutate(sum = sum(numtoday))

month <- c("jan 2020", "feb 2020", "mar 2020", "apr 2020", "may 2020", "jun 2020",
           "jul 2020", "aug 2020", "sep 2020", "oct 2020", "nov 2020", "dec 2020",
           "jan 2021", "")
value <- c(4, 11, 8533, 44688, 37711, 13257, 
           12108, 12636, 29810, 76686, 142695, 203288,
           198426, 87841, 114416, 237308, 162155)
month_num <- c(1:17)

df<-data.frame(month_num, value)

# wave 2 = june 3 - 12
# wave 3 = 2020-10-28 = 
# wave 4 = 2021-01-27 - 2021-02-07
# wave 5 = 2021-03-11 - 2021-03-23
# wave 6 = 2021-05-31 - 2021-06-14
# wave 7 = 2021-09-10 - 2021-09-20

df %>%
  ggplot(aes(x = month_num, y = value)) +
  geom_line() +
  geom_vline(aes(xintercept = 12), alpha = 0.2) +
  geom_vline(aes(xintercept = 13), alpha = 0.2) +
  geom_vline(aes(xintercept = 11), alpha = 0.2) +
  geom_vline(aes(xintercept = 14), alpha = 0.2) +
  geom_vline(aes(xintercept = 10), alpha = 0.2) +
  geom_vline(aes(xintercept = 9), alpha = 0.2) +
  theme_blank()


covid_cases %>%
  filter(date > "2020-06-01") %>% 
  filter(date < "2021-09-30") %>%
  filter(prname == "Canada") %>%
  mutate(group = as.factor(case_when(date > "2020-06-02" & date < "2020-06-13" ~ 1,
                           date > "2020-10-27" & date < "2020-11-11" ~ 2,
                           date > "2021-01-26" & date < "2021-02-08" ~ 3,
                           date > "2021-03-10" & date < "2021-03-24" ~ 4,
                           date > "2021-05-30" & date < "2021-06-15" ~ 5,
                           date > "2021-09-09" & date < "2021-09-21" ~ 6))) %>%
  ggplot(aes(x = date, y = avgtotal_last7)) +
  geom_line(color = "light blue") +
  geom_ribbon(aes(x = date, group = group, ymax = avgtotal_last7, ymin = 0), fill = "light blue", alpha = 0.3) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 month") +
  labs(x = "Time", y = "Number of daily COVID-19 cases")


```