---
title: "CFA: Psychological distress during COVID-19 in Canadian youth with and without chronic health conditions"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load the libraries
library(tidyverse)
library(readr)
library(here)
library(psych)
library(english)
library(stringr)
library(tableone)
library(kableExtra)
library(scales)
library(ggridges)
library(plotly)
library(patchwork)
library(haven)
library(sjPlot)
library(sjmisc)
library(lavaan)
library(stats)
```

```{r data, echo=FALSE, message=FALSE, warning=FALSE}
# Read in the clean csv file

##table(leger_cfa$q0)
# q0 = variable for language

leger_cfa <- read_csv("data/cmaj_analyses.csv") %>%
  filter(wave == 2 | wave == 3) %>%
  dplyr::select(impacvd_sq001, impacvd_sq002, impacvd_sq003, impacvd_sq004) %>%
  drop_na()

leger_cfa_w4 <- read_csv("data/cmaj_analyses.csv") %>%
  filter(wave == 4 | wave == 5) %>%
  dplyr::select(impacvd_sq001, impacvd_sq002, impacvd_sq003, impacvd_sq004) %>%
  drop_na()

## Wave 2 - subdivided for english and french
leger_cfa_w2_en <- read_csv("data/cmaj_analyses.csv") %>%
  filter(wave == 2 | wave == 3) %>%
  filter(q0 == 1) %>%
  dplyr::select(impacvd_sq001, impacvd_sq002, impacvd_sq003, impacvd_sq004) %>%
  drop_na()

leger_cfa_w2_fr <- read_csv("data/cmaj_analyses.csv") %>%
  filter(wave == 2 | wave == 3) %>%
  filter(q0 == 2) %>%
  dplyr::select(impacvd_sq001, impacvd_sq002, impacvd_sq003, impacvd_sq004) %>%
  drop_na()

## Wave 3 - subdivided for english and french
leger_cfa_w4_en <- read_csv("data/cmaj_analyses.csv") %>%
  filter(wave == 4 | wave == 5) %>%
  filter(q0 == 1) %>%
  dplyr::select(impacvd_sq001, impacvd_sq002, impacvd_sq003, impacvd_sq004) %>%
  drop_na()

leger_cfa_w4_fr <- read_csv("data/cmaj_analyses.csv") %>%
  filter(wave == 4 | wave == 5) %>%
  filter(q0 == 2) %>%
  dplyr::select(impacvd_sq001, impacvd_sq002, impacvd_sq003, impacvd_sq004) %>%
  drop_na()
```

#### Aims:

Perform Confirmatory Factor Analyses (CFA) for the psychological distress scale.
In order to examine the reliability of the scale we followed several steps. We examined the factor structure at 2 separate time points (survey round 2 and 3 combined and survey 4&5 combined in order to get a big enough N in each language) along as in English and French. Lastly, we also report Cronbach Alphas.

NB: Survey round equals to survey wave, Leger is the name of the entire dataset and we label waves as wX

**Variable names and description**:

Because of COVID-19...

**impacvd_sq001**	[...I have felt nervous, anxious, or worried]

**impacvd_sq002**	[...I have felt sad, depressed, or hopeless] 

**impacvd_sq003**	[...I have felt lonely and isolated]

**impacvd_sq004**	[...I have felt irritable, frustrated or angry]

The complete list of measures including detailed protocol description, along with the data dictionaries can be found [here](https://osf.io/v8xk5/).

### STEP 1: Examine item to item correlations

##### Zero-order polychoric item to item correlations for wave 2 and 3
```{r echo=FALSE, message = FALSE}
poly_cors<-polychoric(leger_cfa)

items_polychoric <- poly_cors$rho

items_polychoric <- as.data.frame(round(items_polychoric, 2))

items_polychoric[upper.tri(items_polychoric)] <- " "

items_polychoric
  #%>% kable(escape = FALSE) %>% # if rendering html this is more pleasant to look at + interactive :)
#   kable_styling("hover")
```

##### Zero-order polychoric item to item correlations for wave 4 and 5
```{r echo=FALSE, message = FALSE}
poly_cors2<-polychoric(leger_cfa_w4)

items_polychoric2 <- poly_cors2$rho

items_polychoric2 <- as.data.frame(round(items_polychoric2, 2))

items_polychoric2[upper.tri(items_polychoric2)] <- " "

items_polychoric2
# %>%
#   kable(escape = F) %>%
#   kable_styling("hover")
```
### STEP 2: Perform CFA 

CFI stands for Comparative Fit Index, TLI for Tucker-Lewis Index, and RMSEA stands for Root Mean Square Error of Approximation. All are commonly used relative goodness-of-fit indices to assess how well the model fits the data. CFI and TLI values range from 0 to 1, with values => 0.9 indicating acceptable model fit. RMSEA values between 0.05 and 0.08 indicate acceptable fit (Hu & Bentler, 1999).

#### Wave 2 and 3, English scale (*N* = `r nrow(leger_cfa_w2_en)`)

```{r echo=FALSE, message=FALSE}
#one factor three items, default marker method
m1_w2_en  <- ' f  =~ impacvd_sq001 + impacvd_sq002 + impacvd_sq003 + impacvd_sq004'

cfa_m1_w2_en <- cfa(m1_w2_en, data=leger_cfa_w2_en)

summary(cfa_m1_w2_en, fit.measures=TRUE, standardized=TRUE) 
# modindices(cfa_m1, sort. = TRUE)
```

#### Wave 2 abd 3, French scale (*N* = `r nrow(leger_cfa_w2_fr)`)

```{r echo=FALSE, message=FALSE}
#one factor three items, default marker method
m1_w2_fr  <- ' f  =~ impacvd_sq001 + impacvd_sq002 + impacvd_sq003 + impacvd_sq004'

cfa_m1_w2_fr <- cfa(m1_w2_fr, data=leger_cfa_w2_fr)

summary(cfa_m1_w2_fr, fit.measures=TRUE, standardized=TRUE) 
```

#### Wave 4 and 5, English scale (*N* = `r nrow(leger_cfa_w4_en)`)

```{r echo=FALSE, message=FALSE}
#one factor three items, default marker method
m1_w3_en  <- ' f  =~ impacvd_sq001 + impacvd_sq002 + impacvd_sq003 + impacvd_sq004'

cfa_m1_w3_en <- cfa(m1_w3_en, data=leger_cfa_w4_en)

summary(cfa_m1_w3_en, fit.measures=TRUE, standardized=TRUE) 
```

#### Wave 4 and 5, French scale (*N* = `r nrow(leger_cfa_w4_fr)`)

```{r echo=FALSE, message=FALSE}
#one factor three items, default marker method
m1_w3_fr  <- ' f  =~ impacvd_sq001 + impacvd_sq002 + impacvd_sq003 + impacvd_sq004'

cfa_m1_w3_fr <- cfa(m1_w3_fr, data=leger_cfa_w4_fr)

summary(cfa_m1_w3_fr, fit.measures=TRUE, standardized=TRUE) 
```

### STEP 3: Report internal consistency at each time point (Cronbach Alpha)
```{r echo=FALSE, message = FALSE}
leger_w2 <- read_csv("data/cmaj_analyses.csv") %>%
  filter(wave == 2) %>%
  dplyr::select(impacvd_sq001, impacvd_sq002, impacvd_sq003, impacvd_sq004) %>%
  drop_na()
leger_w3 <- read_csv("data/cmaj_analyses.csv") %>%
  filter(wave == 3) %>%
  dplyr::select(impacvd_sq001, impacvd_sq002, impacvd_sq003, impacvd_sq004) %>%
  drop_na()
leger_w4 <- read_csv("data/cmaj_analyses.csv") %>%
  filter(wave == 4) %>%
  dplyr::select(impacvd_sq001, impacvd_sq002, impacvd_sq003, impacvd_sq004) %>%
  drop_na()
leger_w5 <- read_csv("data/cmaj_analyses.csv") %>%
  filter(wave == 5) %>%
  dplyr::select(impacvd_sq001, impacvd_sq002, impacvd_sq003, impacvd_sq004) %>%
  drop_na()
leger_w6 <- read_csv("data/cmaj_analyses.csv") %>%
  filter(wave == 6) %>%
  dplyr::select(impacvd_sq001, impacvd_sq002, impacvd_sq003, impacvd_sq004) %>%
  drop_na()
leger_w7 <- read_csv("data/cmaj_analyses.csv") %>%
  filter(wave == 7) %>%
  dplyr::select(impacvd_sq001, impacvd_sq002, impacvd_sq003, impacvd_sq004) %>%
  drop_na()

psych::alpha(leger_cfa_w2_en) # 0.88
psych::alpha(leger_cfa_w2_fr) # 0.84
psych::alpha(leger_cfa_w4_en) # 0.9
psych::alpha(leger_cfa_w4_fr) # 0.89

psych::alpha(leger_w2) # 0.88
psych::alpha(leger_w3) # 0.87
psych::alpha(leger_w4) # 0.9
psych::alpha(leger_w5) # 0.9
psych::alpha(leger_w6) # 0.89
psych::alpha(leger_w7) # 0.89
```
