---
title: "CMAJ Analytical and results script: Psychological distress during COVID-19 in Canadian youth with and without chronic health conditions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load the libraries
library(tidyverse)
library(readr)
library(here)
library(psych)
library(english)
library(stringr)
library(tableone)
library(kableExtra)
library(scales)
library(ggridges)
library(plotly)
library(patchwork)
library(haven)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(table1)
library(rstatix)
library(car)
library(multcomp)
library(effects)
library(pastecs)
```

```{r data, echo=FALSE, message=FALSE, warning=FALSE}
# Read in the clean csv file
leger <- read_csv("data/cmaj_analyses.csv")

# subset for chronic people only
leger_chronic <- leger %>%
  filter(chronic_dummy == 1)

# Data for the plot
marginal_means <- read_csv("data/marginal_means.csv")

marginal_means_2 <- read_csv("data/marginal_means2.csv")

covid_cases <- read_csv("data/covid19-download.csv")

# set seed for all analyses and wrangling
set.seed(1)
```

#### Aims:
Using Canadian samples of young adults, we aimed to investigate if there exist differences in levels of psychological distress in YAs with physical chronic health conditions, and healthy individuals. In particular we wanted to examine any significant fluctuations in psychological distress (feelings of anxiety, depression, loneliness, irritability) among young people with different medical needs over the course of the pandemic.

## METHOD

The present study is part of a larger International COVID‐19 Awareness and Responses Evaluation (iCARE) study, which is a global multi‐phase research survey conducted through the Montreal Behavioural Medicine Centre. Find more about the study and materials at [www.iCAREstudy.com](www.iCAREstudy.com). 

## RESULTS
### Sample

```{r echo=FALSE, warning=FALSE}
table_data <- leger %>%
  dplyr::select(age_yrs, sex, edu, hoinc, province_full, chronic_descriptive, chronic_dummy, chronic_multi, any_m_health, distress_sum, vaccine, booster_dummy) %>%
  mutate(chronic_dummy = factor(chronic_dummy,
                                levels = c(0:1),
                                labels = c("Healthy", "Chronic Health Condition")),
         chronic_multi = factor(chronic_multi, 
                                levels = c(0:1),
                                labels = c("No", "Yes")),
         any_m_health = factor(any_m_health,
                               levels = c(0:1),
                               labels = c("No", "Yes")))

label(table_data$sex) <- "Sex"
label(table_data$age_yrs) <- "Age"
label(table_data$edu) <- "Education"
label(table_data$province_full) <- "Province"
label(table_data$hoinc) <- "Income"
label(table_data$chronic_descriptive) <- "Chronic illness"
label(table_data$chronic_multi) <- "Multimorbidity"
label(table_data$any_m_health) <- "Diagnosis of mental health disorder"
label(table_data$distress_sum) <- "COVID-19 distress"
label(table_data$vaccine) <- "Received COVID-19 vaccine?"
label(table_data$booster_dummy) <- "Likelihood of taking COVID-19 booster dose?"

units(table_data$age_yrs) <- "years"

table1(~ age_yrs + sex + edu + hoinc + province_full + chronic_descriptive + chronic_multi + any_m_health + distress_sum + vaccine + booster_dummy | chronic_dummy,
       data = table_data)

```

## Within-group comparisons over time, adjusted for age & sex & mental health diagnosis (anxiety or depression)

#### Chronic health conditions over time
```{r echo=FALSE, warning=FALSE, message=FALSE}
chronic <- leger %>%
  dplyr::select(distress_sum, sex, age_yrs, hoinc, chronic_dummy, any_m_health, wave) %>%
  filter(chronic_dummy == 1)%>%
  mutate(wave = factor(wave, ordered = TRUE))

# Levene's test for homogeneity of variance
# levene_test(chronic, distress_sum ~ wave)
##     df1   df2 statistic      p
##   <int> <int>     <dbl>  <dbl>
## 1     5  1448      2.17 0.0553

# ANCOVA
m1 = aov(distress_sum ~ wave + sex + age_yrs + any_m_health, data = chronic)

Anova(m1, type="III")

# To get adjusted or marginal means
# m1_adjusted <- effect("wave", m1, se=TRUE)
# m1_adjusted$se
# summary(m1_adjusted)
# format(2.714e-14, scientific = F, digits = 3)

```

Effect sizes and contrasts for sex, age, mental health
```{r echo=FALSE, warning=FALSE}
# Calculate effect sizes
#

#summary.lm(m1)
# # sex, age, mental health
t_m1 <- c(-7.938, -2.517, 10.138)
df_m1 <- 1438

sqrt((t_m1^2)/(t_m1^2+df_m1)) # <- these are contrasts

# Age - partial eta^2 for the effect of age
61.2/(61.2+13890.6)
```


#### No chronic health conditions over time
```{r echo=FALSE, warning=FALSE, message=FALSE}
healthy <- leger %>%
  dplyr::select(distress_sum, sex, age_yrs, hoinc, chronic_dummy, any_m_health, wave) %>%
  filter(chronic_dummy == 0)%>%
  mutate(wave = factor(wave, ordered = TRUE))

# Leven's test
#levene_test(healthy, distress_sum ~ wave)
##     df1   df2 statistic      p
##   <int> <int>     <dbl>  <dbl>
## 1     5  4399      2.12 0.0605

# ANCOVA
m2 = aov(distress_sum ~ wave + sex + age_yrs + any_m_health, data = healthy)

Anova(m2, type="III")
m2_posthoc <- glht(m2, linfct = mcp(wave = "Tukey"))
summary(m2_posthoc)

# To get adjusted or marginal means
# m2_adjusted <- effect("wave", m2, se=TRUE)
# m2_adjusted$se
# 
# summary(m2_adjusted)


```

Effect sizes in order for sex, age, mental health
```{r echo=FALSE, warning=FALSE}
# Calculate effect sizes

summary.lm(m2)
# # sex, age, mental health, w3, w4, w5, w6, w7
t_m2 <- c(-10.704, 15.372, -1.028, -7.737, 0.947, 0.738, 1.615) # <- contrasts
df_m2 <- 4380

sqrt((t_m2^2)/(t_m2^2+df_m2))
# Partial eta2
# age
904/(904+50374)

## These values were used to calculate cohen d effect size, might have been able to use some function but because was not sure I have done it manually
# Marginal mean and N at w2
6.242078
730
# Marginal mean and N at w3
6.951690
540
# Marginal mean and N at w4
7.104910
608
# Marginal mean and N at w5
7.258187
666

# Calculate the SDs for effect size calculations
se <- c(0.1190545, 0.1333590, 0.1263253, 0.1226668)
N <- c(730, 540, 608, 666)
sd <- se*sqrt(N)
n1 <- 730
s1 <- 3.216675
n2 <-c(540, 608, 666)
s2 <- c(3.098983, 3.114886, 3.165659)

# Get pooled SDs for Cohen d's
pooled <- sqrt(((n1-1)*s1^2 + (n2-1)*s2^2) / (n1+n1-2))
## COHEN's D
# 2-3 Cohen's d = (6.95169 - 6.242078)/2.953614 = 0.2402521
# 2-4 Cohen's d = (7.10491 - 6.242078)/3.035274 = 0.2842682.
# 2-5 Cohen's d = (7.258187 - 6.242078)/3.121586 = 0.3255105
```

### DISTRESS between chronic health condition v. no chronic health at each survey wave -->

#### Wave 2
```{r echo=FALSE}
w2 <- leger %>%
  dplyr::select(distress_sum, sex, age_yrs, any_m_health, chronic_dummy, wave) %>%
  filter(wave == 2)
  
w2_m1 = aov(distress_sum ~ chronic_dummy + age_yrs + any_m_health, data = w2)

Anova(w2_m1, type="III")
```
Effect size = `r 109.9/(109.9+11504.4)`

#### Wave 3
```{r echo=FALSE}
w3 <- leger %>%
  dplyr::select(distress_sum, sex, age_yrs, any_m_health, chronic_dummy, wave) %>%
  filter(wave == 3)

w3_m1 = aov(distress_sum ~ chronic_dummy + age_yrs + any_m_health, data = w3)

Anova(w3_m1, type="III")
```
#### Wave 4
```{r echo=FALSE}
w4 <- leger %>%
  dplyr::select(distress_sum, sex, age_yrs, any_m_health, chronic_dummy, wave) %>%
  filter(wave == 4)

w4_m1 = aov(distress_sum ~ chronic_dummy + age_yrs + any_m_health, data = w4)

Anova(w4_m1, type="III")
```
Effect size = `r 36.9/(36.9+11647.9)`

#### Wave 5
```{r echo=FALSE}
w5 <- leger %>%
  dplyr::select(distress_sum, sex, age_yrs, any_m_health, chronic_dummy, wave) %>%
  filter(wave == 5)

w5_m1 = aov(distress_sum ~ chronic_dummy + age_yrs + any_m_health, data = w5)

Anova(w5_m1, type="III")  
```
#### Wave 6
```{r echo=FALSE}
w6 <- leger %>%
  dplyr::select(distress_sum, sex, age_yrs, any_m_health, chronic_dummy, wave) %>%
  filter(wave == 6)

w6_m1 = aov(distress_sum ~ chronic_dummy + age_yrs +any_m_health, data = w6)

Anova(w6_m1, type="III")   
```
Effect size = `r 68.5/(68.5+10759.7)`

#### Wave 7
```{r echo=FALSE}
w7 <- leger %>%
  dplyr::select(distress_sum, sex, age_yrs, any_m_health, chronic_dummy, wave) %>%
  filter(wave == 7)

w7_m1 = aov(distress_sum ~ chronic_dummy + age_yrs + any_m_health, data = w7)

Anova(w7_m1, type="III")  
```
Effect size = `r 145.8/(145.8+10996.4)`
