---
title: "SENSITIVITY ANALYSES: Comparison of the 2 groups (chronic v. no chronic condition) across waves"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load the libraries
library(tidyverse)
library(readr)
library(here)
library(kableExtra)
library(scales)
library(haven)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(table1)
library(sjstats)
library(tableone)
```

Aim:

This document shows the sensitivity analysis in order to demonstrate that the independent sample across survey rounds are similar enough in terms of demographics so that they can be compared. Because we analyzed independent sub-samples of a representative survey across 6 time points, we compared these samples on the main demographic characteristic within each group: chronic and no chronic health condition. Individuals with chronic health condition did not vary on any variables. Individuals with no chronic health conditions showed minor discrepancies in age, sex, and educational distributions between separate waves. The complete table output is available on the study repository. See below for the exact breakdowns and details.

```{r data, echo=FALSE, message=FALSE, warning=FALSE}
# Read in the clean csv file
leger <- read_csv("data/cmaj_analyses.csv")

leger_chronic <- leger %>%
  filter(chronic_dummy == 1)

leger_healthy <- leger %>%
  filter(chronic_dummy == 0)
```

## TABLE 1: Young adults with chronic health conditions across survey rounds/waves

No significant differences were observed.

```{r echo = FALSE}
## Create a table
# First create a variable list you want in the table

var_list <- c("age_yrs", "sex", "edu", "province_full", "chronic_descriptive", 
              "chronic_multi", "any_m_health", "hoinc")

# Determine which variables are categorical
cat_vars <- c("sex", "edu", "province_full", "chronic_descriptive", 
              "chronic_multi", "any_m_health", "hoinc", "wave")

#All cases table
table_1 <- CreateTableOne(vars = var_list, 
                          factorVars = cat_vars,
                          data = leger_chronic,
                          strata = c("wave"))

table_1
```

## TABLE 2: Young adults with no chronic health conditions across survey rounds/waves

There exist minor differences between some waves in variables age, sex, education level, and perceived income status.

```{r echo = FALSE}
## Create a table
table_2 <- CreateTableOne(vars = var_list, 
                          factorVars = cat_vars,
                          data = leger_healthy,
                          strata = c("wave"))

table_2
```




