---
title: "CPA 2022 abstract"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load the libraries
library(tidyverse)
library(readr)
library(here)
library(psych)
library(english)
library(stringr)
library(tableone)
library(kableExtra)
library(scales)
library(ggridges)
library(plotly)
library(patchwork)
library(haven)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(table1)
library(rstatix)
library(sjstats)
```

```{r data, echo=FALSE, message=FALSE, warning=FALSE}
# Read in the clean csv file
cpa_data <- read_csv("CPA_2021/cpa_analyses.csv")

```
## Weighted descriptive info
* sex distribution F = 51.6% (N = 9,164), M = 49.4% (N = 8,567)
* average age = 48.25 years, SD = 16.9
* depression = yes = 15.9% (N = 2,845), no = 84.1% (N = 14,971)
* anxiety = yes = 21.1% (N = 3,763), no = 78.9% (N = 14,053)
* comorbid anxiety and depression = yes = 12% (N = 2,145), no = 88% (N = 15,671)

```{r echo=FALSE}
library(pollster)
data_table <- cpa_data %>%
  mutate(group = case_when(chronic_dummy == 0 & comorbid_dep_anx == 0 ~ "Healthy",
                           chronic_dummy == 0 & comorbid_dep_anx == 1 ~ "Healthy PD",
                           chronic_dummy == 1 & comorbid_dep_anx == 0 ~ "NCD",
                           chronic_dummy == 1 & comorbid_dep_anx == 1 ~ "NCD PD"))

data_table %>%
  group_by(group) %>%
  count()

weighted_mean(data_table$age_yrs)
weighted_sd(data_table$age_yrs)

topline(df = data_table, variable = comorbid_dep_anx, weight = pond)
```

## OUTCOME 6 - Descriptive Healthy vs healthy PD

I have cancelled medical appointments or going to the emergency room for a non-COVID-related problem
1 = great extent
0 = all others
```{r echo=FALSE, warning=FALSE}
healthy <- cpa_data %>%
  filter(group == 0 | group == 1) %>%
  mutate(group = case_when(group == 0 ~ "Healthy",
                           group == 1 ~ "Healthy PD")) 
healthy %>%
  group_by(group, impacvd_sq006) %>%
  count() %>%
  drop_na %>%
  group_by(group) %>%
  mutate(sum_grp = sum(n),
         prop = round((n/sum_grp*100),2)) %>%
  select(-sum_grp) %>%
  kable() %>%
  kable_styling(full_width = FALSE)

weighted_chisqtest(healthy, 
                   x = impacvd_sq006,
                   y = group,
                   weights = pond)
```

## OUTCOME 6- Descriptive NCD vs NCD PD
```{r echo=FALSE, warning=FALSE}
ncd <- cpa_data %>%
  filter(group == 2 | group == 3) %>%
  mutate(group = case_when(group == 2 ~ "NCD",
                           group == 3 ~ "NCD PD")) 
ncd %>%
  group_by(group, impacvd_sq006) %>%
  count() %>%
  drop_na %>%
  group_by(group) %>%
  mutate(sum_grp = sum(n),
         prop = round((n/sum_grp*100),2)) %>%
  select(-sum_grp) %>%
  kable() %>%
  kable_styling(full_width = FALSE)

weighted_chisqtest(ncd, 
                   x = impacvd_sq006,
                   y = group,
                   weights = pond)
```

## OUTCOME 20 - Descriptive Healthy vs healthy PD
Trouble getting non-COVID-related care
1 = great extent
0 = all others
```{r echo=FALSE, warning=FALSE}
healthy2 <- cpa_data %>%
  filter(group == 0 | group == 1) %>%
  mutate(group = case_when(group == 0 ~ "Healthy",
                           group == 1 ~ "Healthy PD")) 
healthy2 %>%
  group_by(group, impacvd_sq020) %>%
  count() %>%
  drop_na %>%
  group_by(group) %>%
  mutate(sum_grp = sum(n),
         prop = round((n/sum_grp*100),2)) %>%
  select(-sum_grp) %>%
  kable() %>%
  kable_styling(full_width = FALSE)

weighted_chisqtest(healthy2, 
                   x = impacvd_sq020,
                   y = group,
                   weights = pond)
```

## OUTCOME 20 - Descriptive NCD vs NCD PD
```{r echo=FALSE, warning=FALSE}
ncd2 <- cpa_data %>%
  filter(group == 2 | group == 3) %>%
  mutate(group = case_when(group == 2 ~ "NCD",
                           group == 3 ~ "NCD PD")) 

ncd2 %>%
  group_by(group, impacvd_sq020) %>%
  count() %>%
  drop_na %>%
  group_by(group) %>%
  mutate(sum_grp = sum(n),
         prop = round((n/sum_grp*100),2)) %>%
  select(-sum_grp) %>%
  kable() %>%
  kable_styling(full_width = FALSE)

weighted_chisqtest(ncd2, 
                   x = impacvd_sq020,
                   y = group,
                   weights = pond)
```

### LOGIT MODELS - IMP 06 & 20 as outcomes
```{r log. reg, echo=FALSE, warning=FALSE}
# # LOGIT FOR CONCERN 2 - Impact on my health, including dying
cpa_data1 <- cpa_data %>%
  select(impacvd_sq006,impacvd_sq020, group, chronic_dummy, comorbid_dep_anx, age_yrs, sex, wave, pond) %>%
  #mutate(group = as.factor(group)) %>%
  mutate(group = case_when(chronic_dummy == 0 & comorbid_dep_anx == 0 ~ "Healthy",
                           chronic_dummy == 0 & comorbid_dep_anx == 1 ~ "Healthy PD",
                           chronic_dummy == 1 & comorbid_dep_anx == 0 ~ "NCD",
                           chronic_dummy == 1 & comorbid_dep_anx == 1 ~ "NCD PD"))

logit_m1 <- glm(impacvd_sq006 ~ group + age_yrs + sex + wave,
                family = "binomial",
                data = cpa_data1,
                weights = pond)

logit_m2 <- glm(impacvd_sq020 ~ group + age_yrs + sex + wave,
                family = "binomial",
                data = cpa_data1,
                weights = pond)

tab_model(logit_m1, logit_m2)
```

